Binocular combination in the autonomic nervous system

### 2024-06-04 manuscript.tex
Corrected typo in abstract

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{caption} \usepackage{lineno} \linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Binocular combination in the autonomic nervous system},
  colorlinks=true,
  linkcolor={black},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Binocular combination in the autonomic nervous system}
\author{Federico G. Segala\(^1\), Aurelio Bruno\(^{1,2}\), Joel T. Martin\(^{1,3}\), Anisa Y. Morsi\(^1\), Alex R. Wade\(^{1,4}\)\\
\& Daniel H. Baker\(^{1,4}\)}
\date{}

@@

The diameter of the pupil fluctuates in response to levels of ambient light and is regulated by the autonomic nervous system. Increasing light in one eye causes both pupils to constrict, implying the system must combine signals across the two eyes -- a process of binocular integration occuring independently of visual cortex. Distinct classes of retinal photoreceptor are involved in controlling and maintaining pupil diameter, with cones and rods driving the initial constriction and intrinsically photosensitive retinal ganglion cells maintaining diameter over prolonged time periods. Here, we investigate binocular combination by targeting different photoreceptor pathways using the silent substitution method to modulate the input spectra. We find different patterns of binocular response in each pathway, and across the first and second harmonic frequencies. At the first harmonic, luminance and S-cone responses showed strong binocular facilitation, and weak interocular suppression. Melanopsin responses were invariant to the number of eyes stimulated. Notably, the L-M pathway involved binocular inhibition, whereby responses to binocular stimulation were weaker than for monocular stimulation. The second harmonic involved strong interocular suppression in all pathways, but with some evidence of binocular facilitation. Our results are consistent with a computational model of binocular signal combination (implemented in a Bayesian hierarchical framework), in which the weight of interocular suppression differs across pathways. We also find pathway differences in response phase, consistent with different lag times for phototransduction. This work demonstrates for the first time the algorithm governing binocular combination in the autonomic nervous system.

\hypertarget{significance-statement-120-words}{%
\section{Significance statement (120 words)}\label{significance-statement-120-words}}

@@

The autonomic nervous system regulates many involuntary bodily processes, including the constriction and dilation of the pupils in response to light (\protect\hyperlink{ref-McDougal2015}{1}). The anatomical pathway from the retina to the subcortical nuclei controlling the pupillary light response (PLR) is well established: it includes the Pretectal Olivary nucleus (PON), the Superior Cervical ganglion and the Edinger-Westphal nucleus, which project to the iris sphincter muscles that directly control the pupil size (\protect\hyperlink{ref-McDougal2015}{1}--\protect\hyperlink{ref-Wang2015}{3}). Evidence of a binocular component to the PLR is shown by the existence of a consensual response of the pupil (the stimulation of one eye will cause constriction of the other eye; \protect\hyperlink{ref-Wyatt1981}{4}). The anatomical segregation of the subcortical pathway from the rest of the brain means that this binocular combination of signals must occur independently of the cortical processes of binocular integration required for visual perception. Our recent work (\protect\hyperlink{ref-Segala2023}{5}) has shown that the algorithm underlying binocular combination of light in the pupil pathway differs from that in the cortex. Here we extend this paradigm to compare binocular combination in different photoreceptor pathways that feed into the autonomic nervous system.

@@

A technique known as silent substitution (\protect\hyperlink{ref-Estevez1982}{22}) can be used to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them (\protect\hyperlink{ref-Shapiro1996}{23}, \protect\hyperlink{ref-Spitschan2018}{24}, this paper also offers a clear explanation of how to implement silent substitution). Studies using silent substitution have shown that the pupils behave differently depending on which photoreceptor or pathway is being stimulated. For example, pupil responses have been shown to be out of phase when the S cones are stimulated, relative to when the ipRGCs or the L+M cones are stimulated (\protect\hyperlink{ref-Spitschan2014}{10}, \protect\hyperlink{ref-Cao2015}{25}). Additionally, when L and M cones are stimulated individually, there appears to be a paradoxical pupil response: increments in L cone stimulation cause a constriction of the pupil, while increments in M cone stimulation lead to a dilation of the pupil (\protect\hyperlink{ref-Murray2018}{9}, \protect\hyperlink{ref-Woelders2018}{12}), perhaps due to activity in the L-M opponent pathway.

@@

Most studies that use silent substitution to investigate the PLR stimulate only one eye. If the stimulated eye is pharmacologically dilated to control pupil size, the response from the unstimulated eye (known as the consensual response) can be recorded (\protect\hyperlink{ref-Spitschan2014}{10}, e.g. \protect\hyperlink{ref-McDougal2010}{15}). While this approach controls the total amount of light incident at the retina, it does not permit investigation of binocular combination for specific photoreceptor pathways. Here we take a different approach, using a custom binocular multiprimary system (see Figure \ref{fig:spectraplots}a,b) that can modulate each (undilated) eye independently, and a binocular eye tracker to measure the diameter of both pupils. We measure the amplitude of pupil modulation (using Fourier analysis) in response to sinusoidal stimulation at 0.5Hz, for stimuli directed at the luminance pathway, L-M cone pathway, S-cone pathway, and melanopsin pathway (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and between monocular stimulation and a dichoptic masking condition in which the two eyes are stimulated at different frequencies (0.5 and 0.4Hz). The results are interpreted using a contemporary model of binocular vision (\protect\hyperlink{ref-Meese2006}{33}) implemented within a hierarchical Bayesian framework.

@@

In contemporary models of binocular signal combination, the amount of binocular facilitation is determined by the magnitude of interocular suppression, with strong suppression reducing facilitation (\protect\hyperlink{ref-Kingdom2015}{34}). We can estimate the strength of interocular suppression by measuring how much monocular responses are reduced when a dichoptic `mask' is shown to the other eye. In our paradigm, the two components flickered at different frequencies (0.5 and 0.4Hz) so that their responses remained distinct in the Fourier spectrum (e.g. \protect\hyperlink{ref-Busse2009}{35}). The yellow diamond symbols in Figure \ref{fig:FirstHarmonicPlots}a-d show the target responses in this condition, and in most cases were weaker than the monocular responses (red circles). The strongest dichoptic masking is found in the L-M condition, where we also observed the binocular suppression effect. Suppression can also be estimated from the responses at 0.4Hz (Figure \ref{fig:FirstHarmonicPlots}f-i). The reduced amplitude in the binocular cross condition (where the two eyes received different temporal frequencies; purple squares) relative to the 0.4Hz monocular condition (grey circles), and the progressive decline in amplitude of the dichoptic cross response (yellow diamonds) also differ across photoreceptor conditions, showing similar differences to those observed at 0.5Hz.

@@

To present synchronised stimulus modulations independently to each eye, two light engines (SpectraTuneLAB: LEDMOTIVE Technologies, LLC, Barcelona, Spain), each with 10 independently addressable LED colour channels, were integrated into a customised viewing system. The light engines were operated via a Python interface to their REST API (\protect\hyperlink{ref-Martin2022}{39}), which supports synchronous launch and playback of spectral sequences prepared in advance and stored in JSON format. When preparing the spectral sequences, the age of participants was used to account for the yellowing of the lenses. We calculated silent substitution solutions using the \emph{PySilSub} toolbox (\protect\hyperlink{ref-Martin2023}{40}), using linear algebra. The outputs of the two light engines (see Figure \ref{fig:spectraplots}b,c) were calibrated using an Ocean Optics Jaz spectroradiometer, which was wavelength-calibrated to an Argon lamp and intensity calibrated using a NIST-traceable light source. We used the (\protect\hyperlink{ref-Stockman2000}{41}) 10-degree cone fundamentals, and estimates of melanopsin absorbance spectra from CIE S 026 (discussed in \protect\hyperlink{ref-Martin2023}{40}) to calculate \(\alpha\)-opic irradiance.

@@

The diffuser discs were positioned in the objective planes of the lenses of a modified VR headset (SHINECON SC-G01, Dongguan Shinecon Industrial Co.~Ltd., Guangdong, China), which was used by the participants to view the stimuli. The stimuli were two discs of flickering light with a diameter of approximately 30 degrees, which were fused together into a cyclopean percept resembling a donut-shaped ring of light, similar to that used in other studies (\protect\hyperlink{ref-Murray2018}{9}, \protect\hyperlink{ref-Spitschan2014}{10}, e.g., \protect\hyperlink{ref-Barrionuevo2016}{42}, \protect\hyperlink{ref-Spitschan2015}{43}, \protect\hyperlink{ref-Zele2018}{44}). The VR headset modifications allowed for small adjustments to account for individual differences in interpupillary distance and focal length. The use of this set up allowed us to modulate the stimuli in three different ocular configurations, similar to the ones we used in our previous study (\protect\hyperlink{ref-Segala2023}{5}): monocular, binocular and dichoptic. In the monocular configuration, the unstimulated eye still saw a non-flickering disc of mean luminance. A schematic of the stimulation system is shown in Figure \ref{fig:spectraplots}a. Pupillometry data were collected using a binocular Pupil Core eye-tracker headset (Pupil Labs GmbH, Berlin, Germany, \protect\hyperlink{ref-Kassner2014}{45}) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-McDougal2015}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{D. H. McDougal, P. D. Gamlin, \href{https://doi.org/10.1002/cphy.c140014}{Autonomic control of the eye}. \emph{Comprehensive Physiology} \textbf{5}, 439--473 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-McDougal2008}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{D. H. McDougal, P. D. Gamlin, {``Pupillary control pathways.''} in \emph{The Senses: A Comprehensive Reference}, (Elsevier, 2008), pp. 521--536.}

\leavevmode\vadjust pre{\hypertarget{ref-Wang2015}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{C.-A. Wang, D. P. Munoz, \href{https://doi.org/10.1016/j.conb.2015.03.018}{A circuit for pupil orienting responses: Implications for cognitive modulation of pupil size}. \emph{Curr Opin Neurobiol} \textbf{33}, 134--40 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-Wyatt1981}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{H. J. Wyatt, J. F. Musselman, \href{https://doi.org/10.1016/0042-6989(81)90097-3}{Pupillary light reflex in humans: Evidence for an unbalanced pathway from nasal retina, and for signal cancellation in brainstem}. \emph{Vision Res} \textbf{21}, 513--25 (1981).}

\leavevmode\vadjust pre{\hypertarget{ref-Segala2023}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{F. G. Segala, \emph{et al.}, \href{https://doi.org/10.7554/elife.87048}{Different rules for binocular combination of luminance flicker in cortical and subcortical pathways}. \emph{eLife} \textbf{12} (2023).}

\leavevmode\vadjust pre{\hypertarget{ref-Barrionuevo2014}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{P. A. Barrionuevo, \emph{et al.}, \href{https://doi.org/10.1167/iovs.13-13252}{Assessing rod, cone, and melanopsin contributions to human pupil flicker responses}. \emph{Investigative Ophthalmology \& Visual Science} \textbf{55}, 719--27 (2014).}

\leavevmode\vadjust pre{\hypertarget{ref-Barrionuevo2018}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{P. A. Barrionuevo, J. J. McAnany, A. J. Zele, D. Cao, \href{https://doi.org/10.3389/fneur.2018.01140}{Non-linearities in the {Rod} and {Cone} {Photoreceptor} {Inputs} to the {Afferent} {Pupil} {Light} {Response}}. \emph{Frontiers in Neurology} \textbf{9} (2018).}

\leavevmode\vadjust pre{\hypertarget{ref-Dacey2005}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{D. M. Dacey, \emph{et al.}, \href{https://doi.org/10.1038/nature03387}{Melanopsin-expressing ganglion cells in primate retina signal colour and irradiance and project to the {LGN}}. \emph{Nature} \textbf{433}, 749--754 (2005).}

\leavevmode\vadjust pre{\hypertarget{ref-Murray2018}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{I. J. Murray, J. Kremers, D. McKeefry, N. R. A. Parry, \href{https://doi.org/10.1364/JOSAA.35.000B66}{Paradoxical pupil responses to isolated {M}-cone increments}. \emph{Journal of the Optical Society of America A} \textbf{35}, B66--B71 (2018).}

\leavevmode\vadjust pre{\hypertarget{ref-Spitschan2014}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{M. Spitschan, S. Jain, D. H. Brainard, G. K. Aguirre, \href{https://doi.org/10.1073/pnas.1400942111}{Opponent melanopsin and s-cone signals in the human pupillary light response}. \emph{Proceedings of the National Academy of Sciences} \textbf{111}, 15568--15572 (2014).}

\leavevmode\vadjust pre{\hypertarget{ref-Spitschan2019}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{M. Spitschan, C. Cajochen, \href{https://doi.org/10.1111/jpi.12602}{Binocular facilitation in light-mediated melatonin suppression?} \emph{J Pineal Res} \textbf{67}, e12602 (2019).}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Binocular combination in the autonomic nervous system},
  colorlinks=true,
  linkcolor=black,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=Blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{caption} \usepackage{lineno} \linenumbers
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{Binocular combination in the autonomic nervous system}
\author{Federico G. Segala\(^1\), Aurelio Bruno\(^{1,2}\), Joel T. Martin\(^{1,3}\),\\
Anisa Y. Morsi\(^1\), Alex R. Wade\(^{1,4}\) \& Daniel H. Baker\(^{1,4}\)}
\date{}

@@

The diameter of the pupil fluctuates in response to levels of ambient light and is regulated by the autonomic nervous system. Increasing light in one eye causes both pupils to constrict, implying the system must combine signals across the two eyes -- a process of binocular integration occurring independently of visual cortex. Distinct classes of retinal photoreceptor are involved in controlling and maintaining pupil diameter, with cones and rods driving the initial constriction and intrinsically photosensitive retinal ganglion cells maintaining diameter over prolonged time periods. Here, we investigate binocular combination by targeting different photoreceptor pathways using the silent substitution method to modulate the input spectra. We find different patterns of binocular response in each pathway, and across the first and second harmonic frequencies. At the first harmonic, luminance and S-cone responses showed strong binocular facilitation, and weak interocular suppression. Melanopsin responses were invariant to the number of eyes stimulated. Notably, the L-M pathway involved binocular inhibition, whereby responses to binocular stimulation were weaker than for monocular stimulation. The second harmonic involved strong interocular suppression in all pathways, but with some evidence of binocular facilitation. Our results are consistent with a computational model of binocular signal combination (implemented in a Bayesian hierarchical framework), in which the weight of interocular suppression differs across pathways. We also find pathway differences in response phase, consistent with different lag times for phototransduction. This work demonstrates for the first time the algorithm governing binocular combination in the autonomic nervous system.

@@

The autonomic nervous system regulates many involuntary bodily processes, including the constriction and dilation of the pupils in response to light (\protect\hyperlink{ref-McDougal2015}{1}). The anatomical pathway from the retina to the subcortical nuclei controlling the pupillary light response (PLR) is well established: it includes the Pretectal Olivary nucleus (PON), the Superior Cervical ganglion and the Edinger-Westphal nucleus, which project to the iris sphincter muscles that directly control the pupil size (\protect\hyperlink{ref-McDougal2015}{1}--\protect\hyperlink{ref-Wang2015}{3}). Evidence of a binocular component to the PLR is shown by the existence of a consensual response of the pupil (\protect\hyperlink{ref-Wyatt1981}{4}). The anatomical segregation of the subcortical pathway from the rest of the brain means that this binocular combination of signals must occur independently of the cortical processes of binocular integration required for visual perception. Our recent work (\protect\hyperlink{ref-Segala2023}{5}) has shown that the algorithm underlying binocular combination of light in the pupil pathway differs from that in the cortex. Here we extend this paradigm to compare binocular combination in different photoreceptor pathways that feed into the autonomic nervous system.

@@

A technique known as silent substitution (\protect\hyperlink{ref-Estevez1982}{22}) can be used to selectively stimulate specific photoreceptor classes. Silent substitution exploits the fact that each photoreceptor class has a distinct spectral tuning that overlaps with the others. Using a multiprimary system, in which the primaries (i.e.~LEDs) have different spectra, it is possible to target one class of photoreceptors while maintaining the others at a constant activity level, effectively silencing them (\protect\hyperlink{ref-Shapiro1996}{23}, \protect\hyperlink{ref-Spitschan2018}{24}). Studies using silent substitution have shown that the pupils behave differently depending on which photoreceptor or pathway is being stimulated. For example, pupil responses have been shown to be out of phase when the S cones are stimulated, relative to when the ipRGCs or the L+M cones are stimulated (\protect\hyperlink{ref-Spitschan2014}{10}, \protect\hyperlink{ref-Cao2015}{25}). Additionally, when L and M cones are stimulated individually, there appears to be a paradoxical pupil response: increments in L cone stimulation cause a constriction of the pupil, while increments in M cone stimulation lead to a dilation of the pupil (\protect\hyperlink{ref-Murray2018}{9}, \protect\hyperlink{ref-Woelders2018}{12}), perhaps due to activity in the L-M opponent pathway.

@@

Most studies that use silent substitution to investigate the PLR stimulate only one eye. If the stimulated eye is pharmacologically dilated to control pupil size, the response from the unstimulated eye (known as the consensual response) can be recorded (\protect\hyperlink{ref-Spitschan2014}{10}, \protect\hyperlink{ref-McDougal2010}{15}). While this approach controls the total amount of light incident at the retina, it does not permit investigation of binocular combination for specific photoreceptor pathways. Here we take a different approach, using a custom binocular multiprimary system (see Figure \ref{fig:spectraplots}a,b) that can modulate each (undilated) eye independently, and a binocular eye tracker to measure the diameter of both pupils. We measure the amplitude of pupil modulation (using Fourier analysis) in response to sinusoidal stimulation at 0.5Hz, for stimuli directed at the luminance pathway, L-M cone pathway, S-cone pathway, and melanopsin pathway (see Figure \ref{fig:spectraplots}e-l). Our key comparisons are between monocular and binocular stimulation, and between monocular stimulation and a dichoptic masking condition in which the two eyes are stimulated at different frequencies (0.5 and 0.4Hz). The results are interpreted using a contemporary model of binocular vision (\protect\hyperlink{ref-Meese2006}{33}) implemented within a hierarchical Bayesian framework.

@@

In contemporary models of binocular signal combination, the amount of binocular facilitation is determined by the magnitude of interocular suppression, with strong suppression reducing facilitation (\protect\hyperlink{ref-Kingdom2015}{34}). We can estimate the strength of interocular suppression by measuring how much monocular responses are reduced when a dichoptic `mask' is shown to the other eye. In our paradigm, the two components flickered at different frequencies (0.5 and 0.4Hz) so that their responses remained distinct in the Fourier spectrum (\protect\hyperlink{ref-Busse2009}{35}). The yellow diamond symbols in Figure \ref{fig:FirstHarmonicPlots}a-d show the target responses in this condition, and in most cases were weaker than the monocular responses (red circles). The strongest dichoptic masking is found in the L-M condition, where we also observed the binocular suppression effect. Suppression can also be estimated from the responses at 0.4Hz (Figure \ref{fig:FirstHarmonicPlots}f-i). The reduced amplitude in the binocular cross condition (where the two eyes received different temporal frequencies; purple squares) relative to the 0.4Hz monocular condition (grey circles), and the progressive decline in amplitude of the dichoptic cross response (yellow diamonds) also differ across photoreceptor conditions, showing similar differences to those observed at 0.5Hz.

@@

To present synchronised stimulus modulations independently to each eye, two light engines (SpectraTuneLAB: LEDMOTIVE Technologies, LLC, Barcelona, Spain), each with 10 independently addressable LED colour channels, were integrated into a customised viewing system. The light engines were operated via a Python interface to their REST API (\protect\hyperlink{ref-Martin2022}{39}), which supports synchronous launch and playback of spectral sequences prepared in advance and stored in JSON format. When preparing the spectral sequences, the age of participants was used to account for the yellowing of the lenses. We calculated silent substitution solutions using the \emph{PySilSub} toolbox (\protect\hyperlink{ref-Martin2023}{40}), using linear algebra. The outputs of the two light engines (see Figure \ref{fig:spectraplots}b,c) were calibrated using an Ocean Optics Jaz spectroradiometer, which was wavelength-calibrated to an Argon lamp and intensity calibrated using a NIST-traceable light source. We used the (\protect\hyperlink{ref-Stockman2000}{41}) 10-degree cone fundamentals, and estimates of melanopsin absorbance spectra from CIE S 026 (\protect\hyperlink{ref-Martin2023}{40}) to calculate \(\alpha\)-opic irradiance.

@@

The diffuser discs were positioned in the objective planes of the lenses of a modified VR headset (SHINECON SC-G01, Dongguan Shinecon Industrial Co.~Ltd., Guangdong, China), which was used by the participants to view the stimuli. The stimuli were two discs of flickering light with a diameter of approximately 30 degrees, which were fused together into a cyclopean percept resembling a donut-shaped ring of light, similar to that used in other studies (\protect\hyperlink{ref-Murray2018}{9}, \protect\hyperlink{ref-Spitschan2014}{10}, \protect\hyperlink{ref-Barrionuevo2016}{42}--\protect\hyperlink{ref-Zele2018}{44}). The VR headset modifications allowed for small adjustments to account for individual differences in interpupillary distance and focal length. The use of this set up allowed us to modulate the stimuli in three different ocular configurations, similar to the ones we used in our previous study (\protect\hyperlink{ref-Segala2023}{5}): monocular, binocular and dichoptic. In the monocular configuration, the unstimulated eye still saw a non-flickering disc of mean luminance. A schematic of the stimulation system is shown in Figure \ref{fig:spectraplots}a. Pupillometry data were collected using a binocular Pupil Core eye-tracker headset (\protect\hyperlink{ref-Kassner2014}{45}) running at 120 Hz, and the signals were recorded with the Pupil Capture software.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-McDougal2015}{}%
1. D. H. McDougal, P. D. Gamlin, Autonomic control of the eye. \emph{Comprehensive Physiology} \textbf{5}, 439--473 (2015).

\leavevmode\hypertarget{ref-McDougal2008}{}%
2. D. H. McDougal, P. D. Gamlin, ``Pupillary control pathways.'' in \emph{The Senses: A Comprehensive Reference}, (Elsevier, 2008), pp. 521--536.

\leavevmode\hypertarget{ref-Wang2015}{}%
3. C.-A. Wang, D. P. Munoz, A circuit for pupil orienting responses: Implications for cognitive modulation of pupil size. \emph{Curr Opin Neurobiol} \textbf{33}, 134--40 (2015).

\leavevmode\hypertarget{ref-Wyatt1981}{}%
4. H. J. Wyatt, J. F. Musselman, Pupillary light reflex in humans: Evidence for an unbalanced pathway from nasal retina, and for signal cancellation in brainstem. \emph{Vision Res} \textbf{21}, 513--25 (1981).

\leavevmode\hypertarget{ref-Segala2023}{}%
5. F. G. Segala\emph{et al.}, Different rules for binocular combination of luminance flicker in cortical and subcortical pathways. \emph{eLife} \textbf{12} (2023).

\leavevmode\hypertarget{ref-Barrionuevo2014}{}%
6. P. A. Barrionuevo\emph{et al.}, Assessing rod, cone, and melanopsin contributions to human pupil flicker responses. \emph{Investigative Ophthalmology \& Visual Science} \textbf{55}, 719--27 (2014).

\leavevmode\hypertarget{ref-Barrionuevo2018}{}%
7. P. A. Barrionuevo, J. J. McAnany, A. J. Zele, D. Cao, Non-linearities in the Rod and Cone Photoreceptor Inputs to the Afferent Pupil Light Response. \emph{Frontiers in Neurology} \textbf{9} (2018).

\leavevmode\hypertarget{ref-Dacey2005}{}%
8. D. M. Dacey\emph{et al.}, Melanopsin-expressing ganglion cells in primate retina signal colour and irradiance and project to the LGN. \emph{Nature} \textbf{433}, 749--754 (2005).

\leavevmode\hypertarget{ref-Murray2018}{}%
9. I. J. Murray, J. Kremers, D. McKeefry, N. R. A. Parry, Paradoxical pupil responses to isolated M-cone increments. \emph{Journal of the Optical Society of America A} \textbf{35}, B66--B71 (2018).

\leavevmode\hypertarget{ref-Spitschan2014}{}%
10. M. Spitschan, S. Jain, D. H. Brainard, G. K. Aguirre, Opponent melanopsin and s-cone signals in the human pupillary light response. \emph{Proceedings of the National Academy of Sciences} \textbf{111}, 15568--15572 (2014).

\leavevmode\hypertarget{ref-Spitschan2019}{}%
11. M. Spitschan, C. Cajochen, Binocular facilitation in light-mediated melatonin suppression? \emph{J Pineal Res} \textbf{67}, e12602 (2019).

\leavevmode\hypertarget{ref-Woelders2018}{}%
12. T. Woelders\emph{et al.}, Melanopsin- and L-cone--induced pupil constriction is inhibited by S- and M-cones in humans. \emph{Proceedings of the National Academy of Sciences} \textbf{115}, 792--797 (2018).

\leavevmode\hypertarget{ref-Mathot2018}{}%
13. S. Mathôt, Pupillometry: Psychology, physiology, and function. \emph{J Cogn} \textbf{1}, 16 (2018).

\leavevmode\hypertarget{ref-Markwell2010}{}%
14. E. L. Markwell, B. Feigl, A. J. Zele, Intrinsically photosensitive melanopsin retinal ganglion cell contributions to the pupillary light reflex and circadian rhythm. \emph{Clinical and Experimental Optometry} \textbf{93}, 137--149 (2010).

\leavevmode\hypertarget{ref-McDougal2010}{}%
15. D. H. McDougal, P. D. Gamlin, The influence of intrinsically-photosensitive retinal ganglion cells on the spectral sensitivity and response dynamics of the human pupillary light reflex. \emph{Vision Research} \textbf{50}, 72--87 (2010).

\leavevmode\hypertarget{ref-Provencio2000}{}%
16. I. Provencio\emph{et al.}, A Novel Human Opsin in the Inner Retina. \emph{Journal of Neuroscience} \textbf{20}, 600--605 (2000).

\leavevmode\hypertarget{ref-Panda2002}{}%
17. S. Panda\emph{et al.}, Melanopsin (opn4) requirement for normal light-induced circadian phase shifting. \emph{Science} \textbf{298}, 2213--2216 (2002).
